---
title: "NoahCollin_607_Project2"
author: "Noah Collin"
output:
  html_document:
    df_print: paged
---
Noah Collin
607
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```
## DataSet 1: Candy Data
This dataset was posted in the discussion boards on CUNY Blackboards. It was posted by Coffy Andrews-Guo.  This link might not work, but you can see his post here: https://bbhosted.cuny.edu/webapps/discussionboard/do/message?action=list_messages&course_id=_2010110_1&nav=discussion_board&conf_id=_2342995_1&forum_id=_2997791_1&message_id=_53840343_1

Data is from:
https://www.scq.ubc.ca/so-much-candy-data-seriously/

The survey that was actually used is here: 
https://www.scq.ubc.ca/wp-content/uploads/2017/10/candyhierarchysurvey2017.pdf
```{R CandyCSV}
Sys.setenv("VROOM_CONNECTION_SIZE" = 131072 * 20)
rawCandy <-  read.csv("candyhierarchy2017.csv", encoding = "UCS-2LE")
```

```{r CandyCleanup1}

colnames(rawCandy) <-  gsub(("Q\\d+"),"", colnames(rawCandy))

colnames(rawCandy) <-  gsub(("\\.\\.+"),"", colnames(rawCandy))

colnames(rawCandy) <-  gsub(("\\.")," ", 
colnames(rawCandy))

colnames(rawCandy) <- str_trim(colnames(rawCandy), side = "left")

#drop columns that aren't candy questions... Still some questionable "candies" in here...
rawCandy <- rawCandy[(-c(112:110))]
#names(rawCandy)
```

```{r Analysis}
trickOrTreaters <- filter(rawCandy, rawCandy$"GOING OUT" == "Yes")

#OldGrumps <- filter(rawCandy, rawCandy$"GOING OUT" == "No")

##TODO

trickOrTreaters <-  trickOrTreaters %>% rowwise() %>% mutate(SumJoys = sum(c_across(all_of(7:109)) == "JOY"))
trickOrTreaters <-  trickOrTreaters %>% rowwise() %>% mutate(SumDespair = sum(c_across(all_of(7:109)) == "DESPAIR"))

trickOrTreaters <- trickOrTreaters %>% mutate(RatioOfJoyToDespair = SumJoys / sum(SumJoys ,SumDespair))


mostJoyfulTrickOrTreaters <-  arrange(trickOrTreaters, desc(trickOrTreaters$RatioOfJoyToDespair))

mostJoyfulTrickOrTreaters <- mostJoyfulTrickOrTreaters %>% transform()

#class(mostJoyfulTrickOrTreaters$AGE)



mostJoyfulTrickOrTreaters$AGE <-as.numeric( (mostJoyfulTrickOrTreaters$AGE) )

mostJoyfulTrickOrTreaters$AGE <- ifelse(mostJoyfulTrickOrTreaters$AGE < 90, mostJoyfulTrickOrTreaters$AGE, NA)


mostJoyfulTrickOrTreaters %>% ggplot(aes(x=(AGE), y = RatioOfJoyToDespair)) +
  geom_smooth() +
  ggtitle("Age to enjoying various candies while Trick Or treating")
```

40 year olds seem to be the grumpiest trick-or-treaters.  I think this graph partially shows what a silly dataset this is.  I actually looked to see if there was a real relationship for children under 18, but there didn't seem to be one.

## Dataset 2: Ficticious Financial Data
I made a fake dataset here:
```{r making data for second example}
dumb1 <- tibble(Group = c(1,2,3,4,1,2,3,4,1,2,3,4), year = c(2010,2011,2012,2013, 2014,2015,2016,2017, 2018,2019,2020,2021), Qtr.1 = c(12,21,6,2,5,54,23,67,12,111,324,21), Qtr.2 = c(23,5,56,12,61,21,54,99,12,43,69,10),
                Qtr.3 = c(56,12,34,65,87,13,63,58,12,23,22,63),
                Qtr.4 = c(22,33,44,55,87,12,43,54,51,67,12,126))

```

## Gather and seperate
```{r gathering and separating}
cleanTibble <- dumb1 %>% gather(Quarter, Revenue, Qtr.1:Qtr.4)

cleanTibble <-  cleanTibble %>% separate(Quarter, c("Time_Period", "Period_ID"), sep = "\\.")

cleanTibble$year <- as.factor(cleanTibble$year)
```

#Groupby
```{r Revenues_By_Year}
ggplot(cleanTibble, aes(x =year,y = Revenue) ) +
  geom_boxplot() +
  ggtitle("Revenues by year")
```


```{r grouping}

dumbGrouped <- cleanTibble %>% group_by(Period_ID) %>% summarise(mean = mean(Revenue) , n = n()) 

ggplot(dumbGrouped, aes(x=Period_ID, y = mean, colour = Period_ID, fill = Period_ID)) +
  geom_col() +
  ggtitle("Average Quarterly Revenues")

```



## Dataset 3: Fictional Education Outcomes Dataset
I didn't make this dataset.
Source of this fictional test data:
https://www.kaggle.com/spscientist/students-performance-in-exams
also:
http://roycekimmons.com/tools/generated_data/exams

``` {r Tests3}

rawTests <- read_csv("StudentsPerformance.csv")

colnames(rawTests) <-  colnames(rawTests) <-  gsub("\\W","_", colnames(rawTests))
#names(rawTests)
#unique(rawTests$"parental_level_of_education")

rawTests$parental_level_of_education <- factor(rawTests$parental_level_of_education, levels = c("some high school",  "high school", "some college", "associate's degree","bachelor's degree", "master's degree"))

parentalEducationGrouped <- rawTests %>% group_by(parental_level_of_education) %>%
  summarise(meanMath = mean(math_score), meanWriting = mean(writing_score), meanReading = mean(reading_score))

ggplot(parentalEducationGrouped, aes(x = parental_level_of_education, y=meanMath, fill = parental_level_of_education )) +
  geom_col() +
  coord_flip() + 
  ggtitle("Math Scores by Parental Education")


ggplot(parentalEducationGrouped, aes(x = parental_level_of_education, y=meanReading, fill = parental_level_of_education )) +
  geom_col() +
  coord_flip() + 
  ggtitle("Reading Scores by Parental Education")

  ggplot(parentalEducationGrouped, aes(x = parental_level_of_education, y=meanWriting, fill = parental_level_of_education )) +
  geom_col() +
  coord_flip() + 
  ggtitle("Writing Scores by Parental Education")

```

Interestingly, students whose parents have "some high school" seem to outperform those who have a high school diploma in the three R's.